--- 
title: "ASSIGNMENT 4"
author: "Ragunath Gunasekaran"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document: default
  pdf_document: default
bibliography: bibliography.bib
---

# Markdown Basics

## Favorite Foods
1. Pizza
2. Burger
3. Pad Thai

\newpage
## Images


```{r dataframe code1, include=FALSE}
heights_df <- read.csv("heights.csv") 
covid_df <- read.csv("us-states.csv")
covid_df$date <- as.Date(covid_df$date)
covid_df$date <- as.Date(covid_df$date)

california_df <- covid_df[ which( covid_df$state == "California"), ]
ny_df <- covid_df[ which( covid_df$state == "New York"), ]
florida_df <- covid_df[ which( covid_df$state == "Florida"), ]
```

```{r plot1, echo=FALSE, message=FALSE}
library(ggplot2)
ggplot(data=florida_df, aes(x=date, group=1)) +
  geom_line(aes(y = cases, colour = "Florida")) +
  geom_line(data=ny_df, aes(y = cases,colour="New York")) +
  geom_line(data=california_df, aes(y = cases, colour="California")) +
  scale_colour_manual("",
                      breaks = c("Florida", "New York", "California"),
                      values = c("darkred","darkgreen", "steelblue" )) +
  xlab(" ") + ylab("Cases") + scale_y_log10()


```

All Cases (Log Plot)

## Add a Quote
The best way to find yourself is to lose yourself in the service of others

## Add an Equation

$$P(A \mid B)= { P(B \mid A) P(A)}$$ 

## Add a Footnote

This is a footnote [^1].

[^1]: This is a footnote in R Markdown


## Add Citations

* R for Everyone (Lander 2014)
* Discovering Statistics Using R (Field, Miles, and Field 2012)

# Inline Code


## NY Times COVID-19 Data
\newline

```{r plot, echo=FALSE, message=FALSE}
library(ggplot2)
ggplot(data=florida_df, aes(x=sort(date),group=1)) +
  geom_line(aes(y = cases, colour = "Florida")) +
  geom_line(data=ny_df, aes(y = cases,colour="New York")) +
  geom_line(data=california_df, aes(y = cases, colour="California")) +
  scale_colour_manual(name = "State Dataset",,
                      breaks = c("Florida", "New York", "California"),
                      values = c("red", "green", "blue")) +
  xlab(" ") + ylab("Cases")+scale_y_log10()


```

## R4DS Height vs Earnings
\newline

```{r plot2, echo=FALSE, message=FALSE}
ggplot(heights_df, aes(x=height , y=earn, col=sex)) + geom_point()
```

# Tables

## Knitr Table with Kable

```{r ktable, echo=FALSE}
name <- c("Aragon", "Bilbo", "Frodo", "Galadriel", "Sam", "Gandalf", "Legolas", "Sauron", "Gollum")
race <- c("Men", "Hobbit", "Hobbit", "Elf", "Hobbit", "Maia", "Elf", "Maia", "Hobbit")
in_fellowship <- c(TRUE, FALSE, TRUE, FALSE, TRUE, TRUE, TRUE, FALSE, FALSE)
ring_bearer <- c(FALSE, TRUE, TRUE, FALSE, TRUE, TRUE, FALSE, TRUE, TRUE)
age <- c(88, 129, 51, 7000, 36, 2019, 2931, 7052, 589)

characters_df <- data.frame(name, race, in_fellowship, ring_bearer, age)
knitr::kable (head(characters_df[, 1:5]), caption = "One Ring to Rule Them All")
```

\newpage
## Pandoc Table
```{r Pandoctable, echo=FALSE}
library(pander)
#CONVERTING BOLLEAN VALUES TO Y OR N
characters_df$in_fellowship <- ifelse(in_fellowship == TRUE, "Yes", "No")
characters_df$ring_bearer <- ifelse(ring_bearer == TRUE, "Yes", "No")
pandoc.table(characters_df, col.names=c("Name","Race","In Fellowship?","Is Ring Bearer?","age"), style = 'grid')

```



# References
1.  Lander, J. P. 2014. R for Everyone: Advanced Analytics and Graphics. Addison-Wesley Data and Analytics Series. Addison-Wesley. https://books.google.com/books?id=3eBVAgAAQBAJ.

2.  R Core Team. 2020. R: A Language and Environment for Statistical Computing. Vienna, Austria: R Foundation for Statistical Computing. https://www.R-project.org/.

3.  Xie, Yihui. 2016. Bookdown: Authoring Books and Technical Documents with R Markdown. Boca Raton, Florida: Chapman; Hall/CRC. https://github.com/rstudio/bookdown.

4.  https://bookdown.org/yihui/rmarkdown-cookbook/

