--- 
title: "ASSIGNMENT 9 - Exercise 15: Introduction to Machine Learning"
author: "Ragunath Gunasekaran"
date: "`r Sys.Date()`"
output:
  html_document: default
  word_document: default
  pdf_document: default
---


## Footnote
This is a Footnote test.

## Citations

* R for Everyone
* Discovering Statistics Using R


##  a. Plot the data from each dataset using a scatter plot

```{r dataframe datasets, include=FALSE}
setwd("C:/Users/ragun/Documents/GitHub/dsc520-master/DSC520-new")
# The first dataset (found in binary-classifier-data.csv) contains three variables; label, x, and y. 
#The label variable is either 0 or 1 and is the output we want to predict using the x and y variables. 
bin_class_data <- read.csv("data/binary-classifier-data.csv")

#The second dataset (found in trinary-classifier-data.csv) is similar to the first dataset except that the label variable can be 0, 1, or 2.
tri_class_data <- read.csv("data/trinary-classifier-data.csv")
```


Binary classifier Data - Scatter plot displays the relationship between x and y variables

```{r Scatterbinclassplot, echo=FALSE, message=FALSE}
library(ggplot2)

scatter_bin_class_data <- ggplot(bin_class_data, aes(x, y))
# Adding geom_point and Linear Regression line with considering Labels
scatter_bin_class_plot <- scatter_bin_class_data + geom_point() + geom_smooth(method = "lm", aes(fill = label), alpha = 0.1) + labs(x = "Value of X", y = "Value of y", colour = "label")
#plot the results - binary classifier dataset
plot(scatter_bin_class_plot)

```


Trinary classifier Data -  Scatter plot displays the relationship between x and y variables


```{r Scattertriclassplot, echo=FALSE, message=FALSE}

scatter_tri_class_data <- ggplot(tri_class_data, aes(x, y))
# Adding geom_point and Linear Linear Regression line with considering Labels
scatter_tri_class_plot <- scatter_tri_class_data + geom_point() + geom_smooth(method = "lm", aes(fill = label), alpha = 0.1) + labs(x = "Value of X", y = "Value of y", colour = "label")
#plot the results - trinary classifier dataset
plot(scatter_tri_class_plot)

```


## Fitting a model is when you use the input data to create a predictive model.Accuracy is simply the percentage of how often the model predicts the correct result. If the model always predicts the correct result, it is 100% accurate. If the model always predicts the incorrect result, it is 0% accurate.


```{r dataframe databintriknn, include=FALSE}
# Data Normalization
binary_data_nor <- bin_class_data[, c("x", "y")]
trinary_data_nor <- tri_class_data[, c("x", "y")]

# Create train and test datasets for binary data
set.seed(111)
dat.binarydataset <- sample(1:nrow(binary_data_nor),size=nrow(binary_data_nor)*0.7,replace = FALSE) #random selection of 70% data.
 
train.binarydataset <- bin_class_data[dat.binarydataset,] # 70% training data
test.binarydataset <- bin_class_data[-dat.binarydataset,] # remaining 30% test data


# Create train and test datasets for trinary data
set.seed(123)
dat.trinarydataset <- sample(1:nrow(trinary_data_nor),size=nrow(trinary_data_nor)*0.7,replace = FALSE) #random selection of 70% data.
 
train.trinarydataset <- tri_class_data[dat.trinarydataset,] # 70% training data
test.trinarydataset <- tri_class_data[-dat.trinarydataset,] # remaining 30% test data


train.binarydatasetlabels <- bin_class_data[dat.binarydataset,1,drop=TRUE]
test.binarydatasetlabels <-bin_class_data[-dat.binarydataset,1,drop=TRUE]

#Creating seperate dataframe for trinary label data
train.trinarydatasetlabels <- tri_class_data[dat.trinarydataset,1,drop=TRUE]
test.trinarydatasetlabels <-tri_class_data[-dat.trinarydataset,1,drop=TRUE]


```

Number of observations for Binary Dataset - Train, Test

```{r observationsbin, echo=FALSE, message=FALSE}
##  Find the number of observations in binary train data
NROW(train.binarydataset) 

#  Find the number of observations in binary test data
NROW(test.binarydataset)

#  Find the number of observations in binary train data
NROW(train.binarydatasetlabels)

#  Find the number of observations in binary test data
NROW(test.binarydatasetlabels)

```

Number of observations for Binary Dataset - Train, Test

```{r observationstri, echo=FALSE, message=FALSE}

# Find the number of observations in trinary train data
NROW(train.trinarydataset) 

#  Find the number of observations in trinary test data
NROW(test.trinarydataset)

##  Find the number of observations in trinary train data
NROW(train.trinarydatasetlabels) 

#  Find the number of observations in trinary test data
NROW(test.trinarydatasetlabels)

```

## b. Fiting a k nearest neighbors model for each dataset for k=3, k=5, k=10, k=15, k=20, and k=25. Compute the accuracy of the resulting models for each value of k. Plot the results in a graph where the x-axis is the different values of k and the y-axis is the accuracy of the model.

Binary Dataset - k nearest neighbors model

```{r plotbintriknnbin, echo=FALSE, message=FALSE}
#Build models with different K values for binary data
#Store K values and accuracy values in a dataframe for different k values
library(class)
list2 <- list(3,5,10,15,20,25)
i=1
k.binaryaccurcy=1
for (i in list2){
    knn.mod <- knn(train=train.binarydataset, test=test.binarydataset, cl=train.binarydatasetlabels, k=i)
    k.binaryaccurcy[i] <- 100 * sum(test.binarydatasetlabels == knn.mod)/NROW(test.binarydatasetlabels)
    k=i
    cat(k,'=',k.binaryaccurcy[i],'')
    }
    
#  Plot Accuracy values with k values
plot(k.binaryaccurcy, type="b", xlab="K- Value",ylab="Accuracy level in YAxis")

```

Trinary Dataset - k nearest neighbors model

```{r plotbintriknntri, echo=FALSE, message=FALSE}
#Store K values and accuracy values in a dataframe for different k values
library(class)
list3 <- list(3,5,10,15,20,25)
j=1
k.trinaryaccurcy=1
for (j in list3){
    knn.modtri <- knn(train=train.trinarydataset, test=test.trinarydataset, cl=train.trinarydatasetlabels, k=j)
    k.trinaryaccurcy[j] <- 100 * sum(test.trinarydatasetlabels == knn.modtri)/NROW(test.trinarydatasetlabels)
    k=j
    cat(k,'=',k.trinaryaccurcy[j],'')
    }
    
    


#  Plot Accuracy values with k values
plot(k.trinaryaccurcy, type="b", xlab="K- Value in XAxis",ylab="Accuracy level in YAxis")
```

## c. Looking back at the plots of the data, do you think a linear classifier would work well on these datasets?

By looking at the plots, I don't think Linear classifier would work well on these datasets since not seeing the linear combination of the characteristics.

( As per Ref 5, Linear classifier achieves this by making a classification decision based on the value of a linear combination of the characteristics )

# References
1. Bernard Marr. (2016). Supervised V Unsupervised Machine Learning – What’s The Difference?
2. Bernard Marr. (2016). What Is The Difference Between Artificial Intelligence And Machine Learning?
3. Bernard Marr. (2016). What Is The Difference Between Deep Learning, Machine Learning and AI?
4. http://rstudio-pubs-static.s3.amazonaws.com/515845_ac9b3e6ee0b14ede809e2ce59dd43830.html
5. https://en.wikipedia.org/wiki/Linear_classifier